---
title: "DDMO Project"
author: "Group 2"
date: "08/03/2021"
output: html_document
version: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Important settings:

```{r}
Sys.setenv("CUDA_VISIBLE_DEVICES" = -1)
# Number of loops
maxi=1 # if max time budget is used, should be =1

# File name where to save and load the results
filename<-'results/00X1234.Rdata'

# Budget for tuning in seconds (max 15*3600)
timebudget <- 15*3600 
```


Now, you can proceed Part 1 to conduct experiments, or start from Part 2 to proceed Rdata file (line 207). 

-----------------------------------------------------
# Part 1. Conduct the experiment
-----------------------------------------------------

Data Set: the Census-Income (KDD) Data Set
```{r}
## devtools::install_gitlab("bartz/spot",host="http://owos.gm.fh-koeln.de:8055")

library("OpenML")
library("mlr")
library("plotly")
library("SPOT")
library("farff")

if (packageVersion("SPOT") < "2.4.8") message("Please update 'SPOT'")

dirName= "oml.cache"
if (! (dir.exists(dirName))) dir.create(dirName)
setOMLConfig(apikey = "c1994bdb7ecb3c6f3c8f3b35f4b47f1f",
              cachedir = dirName # default cachedir is only temporary.
 )
 dataOML <- getOMLDataSet(4535)$data
 
 # Different input data
 #newdataOML<-dataOML[20001:299285,] 
 #dataOML<-newdataOML
 
 task <- makeClassifTask(data=dataOML,target="V42")
 
 rsmpl <- makeResampleDesc("Holdout", split=0.6)
```

Project Setup
```{r}
task.type <- "classif"
data.seed <- 1
tuner.seed <- 1
timeout <- timebudget/20 # use 1/20 times the budget before tuning is stopped
```

Learner (Algorithm) Definition
```{r}
model <- "xgboost"
learner <- paste(task.type, model, sep=".")
```

Experiment Configuration
```{r}
if(model=="xgboost"){
   tunepars <- c("nrounds","eta","lambda","alpha","subsample","colsample_bytree","gamma","max_depth","min_child_weight")
   lower <- c( 1,  0.4,  -4, -4,  0.8,   0.2,  -4,   3,  -4)
   upper <- c(200,   1,   9,  7,    1,   0.8,   4,  15,   7)
   type <-  c("integer","numeric","numeric","numeric","numeric","numeric","numeric","integer","numeric")
   if(task.type=="classif"){
      fixpars <- list(eval_metric="logloss",# suppress warning given when default metric is used.
                      nthread=1) #one thread, not parallel
      }else{
         fixpars <- list(eval_metric="rmse",# suppress warning given when default metric is used.
                         nthread=1) #one thread, not parallel
         }
   factorlevels <- list()
   transformations <- c(trans_id, trans_id, trans_2pow, trans_2pow, trans_id, trans_id, trans_2pow, trans_id, trans_2pow)
   dummy=TRUE
   relpars <- list()
   }
```

Input missing values
```{r}
task <- impute(
  task,
  classes = list(
    factor = imputeMode(),
    integer = imputeMedian(),
    numeric = imputeMean()
    )
  )$task
```

Replace all factor features with their dummy variables. Internally model.matrix is used. Non factor
features will be left untouched and passed to the result.
```{r}
if(dummy){
  task <- createDummyFeatures(task)
}
# str(task)
```

Set Seed
```{r}
set.seed(data.seed)
```

Compile the information as a list.
```{r}
cfg <- list(
 learner=learner,
   tunepars=tunepars,
   lower=lower,
   upper=upper,
   type=type,
   fixpars=fixpars,
   factorlevels=factorlevels,
   transformations=transformations,
   dummy=dummy,
   relpars=relpars,
   task=task,
   resample = rsmpl
)
```

Get objective function
```{r}
objf <- get_objf(config=cfg, timeout=timeout)
```

Launch SPOT in loops
*maxi* represents number of loops (is assigned at the beginning)

```{r}
#initialize result vectors
ybest_list<- NULL
xbest_list<- list(NULL)
result_list<- list(NULL)

for (index in 1:maxi) {
   ybest_list[index] <- NULL
   xbest_list[index] <- list(NULL)
   result_list[index]<- list(NULL)
}

for (index in 1:maxi) {
   # Your Tuning Run with SPOT
   set.seed(2*index)  # seed is changed
   result <- spot(fun = objf,
                 lower=cfg$lower,
                 upper=cfg$upper,
                 control = list(types=cfg$type,
                                maxTime = timebudget/60, #convert to minutes
                                plots=TRUE,
                                progress = TRUE,
                                model=buildKriging,
                                optimizer=optimDE,
                                noise=TRUE,
                                seedFun=index, # seed is changed  
                                seedSPOT=tuner.seed+index, # seed is changed 
                                designControl=list(size=5*length(cfg$lower)),
                                funEvals=Inf,
                                modelControl=list(target="y",
                                                  useLambda=TRUE,
                                                  reinterpolate=TRUE),
                                optimizerControl=list(funEvals=100*length(cfg$lower))
                 )
   )
   ybest_list[index]<-result$ybest
   xbest_list[[index]]<-result$xbest
   result_list[[index]]<-result
}
```

Save result of experiment
```{r}
save(result_list, file=filename)
```

Save the input of experiment (all configurations).
It will be saved in the same folder as specified at the beginning.
```{r}
input<-list(task.type, data.seed, tuner.seed,timebudget,timeout,model,learner,lower,upper,type,fixpars,factorlevels,transformations,dummy,relpars)

names(input)<-c('task.type','data.seed','tuner.seed','timebudget','timeout','model','learner','lower','upper','type','fixpars','factorlevels','transformations','dummy','relpars')

save(input, file=paste(dirname(filename),'/input_',basename(filename),sep=''))
```

-----------------------------------------------------
# Part 2. Data analyzation
-----------------------------------------------------

Load result
```{r}
load(filename)
```

It is possible to check some experiment configuration in input_file
```{r}
load(paste(dirname(filename),'/input_',basename(filename),sep=''))

# For example, time budget
input$timebudget
```
If the experiment have not conducted in current session (you have directly loaded the R.data file), please, proceed the next chunk. 
```{r}
# Load libraries
library("plotly")
library("SPOT")

maxi<-length(result_list)

# Initialize empty lists
ybest_list<- NULL
xbest_list<- list(NULL)

for (index in 1:maxi) {
   ybest_list[index] <- NULL
   xbest_list[index] <- list(NULL)
}

for (index in 1:maxi) {
   ybest_list[index]<-result_list[[index]]$ybest
   xbest_list[[index]]<-result_list[[index]]$xbest
}
```

Object function results were found
```{r}
ybest_list
```

Hyperparamters that were found
```{r}
xbest_list
```

Make a list for each hyperparameter (separate list for each hyperparameter)

```{r}
nrounds_list<- NULL
eta_list<- NULL
lambda_list<- NULL
alpha_list<- NULL
subsample_list<- NULL
colsample_bytree_list<- NULL
gamma_list<- NULL
max_depth_list<- NULL
min_child_weight_list<- NULL

for (index in 1:maxi) {
   nrounds_list[index] <- xbest_list[[index]][1]
   eta_list[index] <- xbest_list[[index]][2]
   lambda_list[index] <- xbest_list[[index]][3]
   alpha_list[index] <- xbest_list[[index]][4]
   subsample_list[index] <- xbest_list[[index]][5]
   colsample_bytree_list[index] <- xbest_list[[index]][6]
   gamma_list[index] <- xbest_list[[index]][7]
   max_depth_list[index] <- xbest_list[[index]][8]
   min_child_weight_list[index] <- xbest_list[[index]][9]
}

# Now we have a list, that containts results of hyperparameter from all experiments
nrounds_list
```

Parallel plot for single experiment

```{r}
# Choose the number of experiment: from 1 to maxi
experiment_num=1

plot_parallel(result_list[[experiment_num]], yvar = 1, xlab = c("nrounds","eta","lambda","alpha","subsample","colsample_bytree","gamma","max_depth","min_child_weight"))
```
```{r, eval=FALSE}
plot_parallel(result_list[[1]], yvar = 1, xlab = c("nrounds","eta","lambda","alpha","subsample","colsample_bytree","gamma","max_depth","min_child_weight"))
plot_parallel(result_list[[2]], yvar = 1, xlab = c("nrounds","eta","lambda","alpha","subsample","colsample_bytree","gamma","max_depth","min_child_weight"))
plot_parallel(result_list[[3]], yvar = 1, xlab = c("nrounds","eta","lambda","alpha","subsample","colsample_bytree","gamma","max_depth","min_child_weight"))
plot_parallel(result_list[[4]], yvar = 1, xlab = c("nrounds","eta","lambda","alpha","subsample","colsample_bytree","gamma","max_depth","min_child_weight"))
plot_parallel(result_list[[5]], yvar = 1, xlab = c("nrounds","eta","lambda","alpha","subsample","colsample_bytree","gamma","max_depth","min_child_weight"))
```




Plot surface
-> Experiment number can be changed
-> Hyperparameters combination can be changed

```{r}
# Change the number of experiment: from 1 to maxi
experiment_num=1

# Make plots
p1<-plot_surface(result_list[[experiment_num]], which=c(1,2), yvar=1, ylab = "nrounds - eta")
p2<-plot_surface(result_list[[experiment_num]], which=c(3,4), yvar=1, ylab = "lambda - alpha")
p3<-plot_surface(result_list[[experiment_num]], which=c(5,6), yvar=1, ylab ="subsample - colsample_bytree")
p4<-plot_surface(result_list[[experiment_num]], which=c(7,8), yvar=1, ylab ="gamma - max_depth")
p5<-plot_surface(result_list[[experiment_num]], which=c(8,9), yvar=1,  ylab = "max_depth - min_child_weight")

# Title string
plot_title<-paste('Experiment no. ',toString(experiment_num),' / ',toString(maxi),' from file: ',basename(filename))

# Merge several plots
fig <- subplot(p1, p2,p3,p4, nrows=2)
fig <- fig %>% layout(
    title = plot_title)
fig

# Make single plot
plot_surface(result_list[[experiment_num]], which=c(1,2), yvar=1, ylab = plot_title, xlab=c('nrounds','eta'))
```

Create Scatterplot matrices. It is like 'pairs', but interactive. We can use options "Box select" or "Lasso select" to highlight specific points.
[https://plotly-r.com/arranging-views.html#arranging-plotly-objects]

```{r}
# Create data frame
df<-do.call(rbind, Map(data.frame, 
                       nrounds=nrounds_list, 
                       eta=eta_list, 
                       lambda=lambda_list,
                       alpha=alpha_list,
                       # subsample=subsample_list,
                       # colsample_bytree=colsample_bytree_list,
                       # gamma=gamma_list,
                       # max_depth=max_depth_list,
                       min_child_weight=min_child_weight_list,
                       ybest=ybest_list))

# Crate interactive plots

dims <- dplyr::select_if(df, is.numeric)
dims <- purrr::map2(dims, names(dims), ~list(values=.x, label=.y))
plot_ly(
  type = "splom", dimensions = setNames(dims, NULL),
  showupperhalf = FALSE, diagonal = list(visible = FALSE)
)
```

